---
title: "ASLeaf Bacteria 2022"
author: "JReceveur"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=10)
knitr::opts_chunk$set(fig.align="center")
```




# Import{.tabset}

## Sample Overview

```{r,warning=F,include=F}
#From https://gist.github.com/grabear/018e86413b19b62a6bb8e72a9adba349
#Parse Silva function

parse_taxonomy_silva_138 <- function(char.vec){
  # Use default to assign names to elements in case problem with greengenes prefix
  char.vec = parse_taxonomy_default(char.vec)
  # Check for unassigned taxa
  if (char.vec["Rank1"] == "Unassigned") {
    char.vec <- c(Rank1="d__Unassigned", Rank2="d__Unassigned", Rank3="d__Unassigned", Rank4="d__Unassigned",
                  Rank5="d__Unassigned", Rank6="d__Unassigned", Rank7="d__Unassigned")
  }
  # Define the meaning of each prefix according to SILVA taxonomy
  Tranks = c(Rank1="Kingdom", Rank2="Phylum", Rank3="Class", Rank4="Order", Rank5="Family", Rank6="Genus", Rank7="Species")
  # Check for prefix using regexp, warn if there were none. trim indices, ti
  ti = grep("[[:alpha:]]\\_\\_", char.vec)
  if( length(ti) == 0L ){
    warning(
      "No silva prefixes were found. \n",
      "Consider using parse_taxonomy_delfault() instead if true for all OTUs. \n",
      "Dummy ranks may be included among taxonomic ranks now."
    )
    # Will want to return without further modifying char.vec
    taxvec = char.vec
    # Replace names of taxvec according to prefix, if any present...
  } else {
    # Format character vectors for Ambiguous taxa
    if( length(ti) < 7 ){
      for (key in names(char.vec)) {
        if ( char.vec[key] == "Ambiguous_taxa" ) {
          tax_no <- (as.numeric(substr(key, 5, 5)) - 1)
          char.vec[key] = sprintf("d__Ambiguous_taxa", tax_no)
        }
      }
      # Reset the trimmed indicies if Ambiguous taxa
      ti = grep("[[:alpha:]]\\_\\_", char.vec)
    }
    # Remove prefix using sub-"" regexp, call result taxvec
    taxvec = gsub("[[:alpha:]]\\_\\_", "", char.vec)
    # Define the ranks that will be replaced
    repranks = Tranks[substr(char.vec[ti], 1, 3)]
    # Replace, being sure to avoid prefixes notK present in Tranks
    names(taxvec)[ti[!is.na(repranks)]] = repranks[!is.na(repranks)]
  }
  return(taxvec)
}




library(vegan)
library(ggplot2)
library(plyr)
library(dplyr)
library(phyloseq)
library(randomForest)
library(knitr)




set.seed(10)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E484", "#0072B2", "#D55E00", "#000000","#CC79A7") #Build a colorblind palette



theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme

biom=import_biom("ASLeafDec2020_7k.biom",parseFunction=parse_taxonomy_silva_138) #Import biom file which contains taxa names and abundances 
colnames(tax_table(biom))<-c(Rank1="Kingdom", Rank2="Phylum", Rank3="Class", Rank4="Order", Rank5="Family", Rank6="Genus", Rank7="Species")




metadata=read.table("LauraMetadata9Dec2020.txt",header = TRUE)

metadata$Date = factor(metadata$Date, levels = c("3_1","24_1","14_2","7_3","29_3","19_4")) #fixes x-axis labels
metadata$Paired<-paste0(metadata$Site,metadata$Date)

tree=read_tree("ASLeafTreeDec2020.nwk")

sampdat=sample_data(metadata)
sample_names(sampdat)=metadata$ID # If using fungal data change ID to FungalID
physeq=merge_phyloseq(biom,sampdat,tree)
physeq<-filter_taxa(physeq, function (x) {sum(x > 0) > 1}, prune=TRUE)#Remove singletons and zeros
physeq<-subset_samples(physeq, Site!="<NA>") # Remove Control
physeq
#physeq 



GPr  = transform_sample_counts(physeq, function(x) x / sum(x) ) #transform samples based on relative abundance
PhylumAll=tax_glom(GPr, "Phylum")

PhylumLevel = filter_taxa(PhylumAll, function(x) mean(x) > 3e-2, TRUE) #filter out any taxa lower than 3%
FamilyAll=tax_glom(GPr,"Family")
FamilyLevel = filter_taxa(FamilyAll, function(x) mean(x) > 3e-2, TRUE) #filter out any taxa lower than 3%
GenusAll=tax_glom(GPr,"Genus")
GenusLevel = filter_taxa(GenusAll, function(x) mean(x) > 2e-2, TRUE) #filter out any taxa lower than 2%

```


Unique bacterial ASVs between leaf types
```{r}
df <- psmelt(physeq)
head(df)
length(unique(df$OTU))

OakSubset<-subset(df,Treatment=="Q")
OakSubset<-subset(OakSubset,Abundance!=0)
length(unique(OakSubset$OTU))

ChestnutSubset<-subset(df,Treatment=="C")
ChestnutSubset<-subset(ChestnutSubset,Abundance!=0)
length(unique(ChestnutSubset$OTU))

length(intersect(OakSubset$OTU,ChestnutSubset$OTU)) #2228


length(setdiff(ChestnutSubset$OTU,OakSubset$OTU)) #311 ASVs in Chestnut but not in Oak
length(setdiff(OakSubset$OTU,ChestnutSubset$OTU)) # 454 ASVs in Oak but not in Chestnut


## Unique ASVs by day

#########
# Day 21
#########

Timepoint1<-subset_samples(physeq,Days==21)
df1 <- psmelt(Timepoint1)

OakSubset1<-subset(df1,Treatment=="Q")
OakSubset1<-subset(OakSubset1,Abundance!=0)
length(unique(OakSubset1$OTU)) #1106

ChestnutSubset1<-subset(df1,Treatment=="C")
ChestnutSubset1<-subset(ChestnutSubset1,Abundance!=0)
length(unique(ChestnutSubset1$OTU)) #1108

length(intersect(OakSubset1$OTU,ChestnutSubset1$OTU)) #594
length(union(OakSubset1$OTU,ChestnutSubset1$OTU)) #1520

length(setdiff(ChestnutSubset1$OTU,OakSubset1$OTU)) #414 ASVs in Chestnut but not in Oak
length(setdiff(OakSubset1$OTU,ChestnutSubset1$OTU)) # 512 ASVs in Oak but not in Chestnut

#########
# Day 126
#########

Timepoint126<-subset_samples(physeq,Days==126)
df126 <- psmelt(Timepoint126)

OakSubset126<-subset(df126,Treatment=="Q")
OakSubset126<-subset(OakSubset126,Abundance!=0)
length(unique(OakSubset126$OTU)) #1530

ChestnutSubset126<-subset(df126,Treatment=="C")
ChestnutSubset126<-subset(ChestnutSubset126,Abundance!=0)
length(unique(ChestnutSubset126$OTU)) #1534

length(intersect(OakSubset126$OTU,ChestnutSubset126$OTU)) #1095
length(union(OakSubset126$OTU,ChestnutSubset126$OTU)) #1969

length(setdiff(ChestnutSubset126$OTU,OakSubset126$OTU)) #439 ASVs in Chestnut but not in Oak at day 126
length(setdiff(OakSubset126$OTU,ChestnutSubset126$OTU)) # 435 ASVs in Oak but not in Chestnut at day 126

```




# Bacterial Shannon Diversity

```{r}
Shannon<-plot_richness(physeq, x="Days",color="RiverType",shape="Treatment", measures=c("Shannon"))+ylab("Shannon Diversity")+geom_point()+scale_fill_manual(cbPalette)+geom_point(size=5)+facet_wrap(~Site)
Shannon
ShannonData<-Shannon$data


ShannonAOV<-aov(value~RiverType*Treatment*as.factor(Days),data=ShannonData)
summary(ShannonAOV)
# plot(ShannonAOV) #plots ok at best, no major bias by fitted values
hist(resid(ShannonAOV))
shapiro.test(resid(ShannonAOV)) # Shapiro test p = 0.065
library(ggpubr)
ggpaired(ShannonData,x="Treatment2",y="value",id="Paired",color="Treatment",line.color = "gray", line.size = 0.4,)+ylab("Bacterial Shannon Diversity")+xlab("Leaf Type")+stat_compare_means(paired=T)+rremove("legend")


CData<-subset(ShannonData,Treatment=="C")
QData<-subset(ShannonData,Treatment=="Q")


CShannonAOV<-aov(value~RiverType*as.factor(Days),data=CData)
summary(CShannonAOV)
hist(resid(CShannonAOV))
plot(CShannonAOV)
shapiro.test(resid(CShannonAOV)) #p=0.45

QShannonAOV<-aov(value~RiverType*as.factor(Days),data=QData)
summary(QShannonAOV)
hist(resid(QShannonAOV))
plot(QShannonAOV)
shapiro.test(resid(QShannonAOV)) #p=0.078



kruskal.test(CData$value, g=CData$RiverType)
kruskal.test(QData$value, g=QData$RiverType)



compare_means(value ~ RiverType, data = CData, group.by = "Date", p.adjust.method = "fdr",method="kruskal.test") # ns
compare_means(value ~ RiverType, data = QData, group.by = "Date", p.adjust.method = "fdr",method="kruskal.test") # ns



kruskal.test(CData$value, g= CData$Date)
kruskal.test(QData$value, g= QData$Date)


CDataPer<-subset(CData, RiverType=="Perennial")
kruskal.test(CDataPer$value, g=CDataPer$Date)
QDataPer<-subset(QData, RiverType=="Perennial")
kruskal.test(QDataPer$value, g=QDataPer$Date)



# compare_means(
#   value~Treatment,
#   ShannonData,
#   method = "wilcox.test",
#   paired = T,
#   id = "Paired",
#   ref.group = "C",
#   p.adjust.method = "fdr"
# )
# 

CombinedDifference<-data.frame()
for (i in unique(ShannonData$Paired)){
  Subset<-subset(ShannonData, Paired==i)
  LeafQ<-subset(Subset,Treatment=="Q")
  LeafC<-subset(Subset,Treatment=="C")
  Difference<-LeafQ$value-LeafC$value
  CombinedDifference<-rbind(CombinedDifference,Difference)
}
colnames(CombinedDifference)<-"Difference"
hist(CombinedDifference$Difference)
mean(CombinedDifference$Difference)
sd(CombinedDifference$Difference)/sqrt(length(CombinedDifference$Difference))



ShannonSummary <- ddply(ShannonData, c("Treatment2","Days","RiverType"), summarise,
                 N    = length(value),
                 mean = mean(value),
                 sd   = sd(value),
                 se   = sd / sqrt(N)
)

BacterialShannonPlot<-ggplot(ShannonSummary,aes(x=as.factor(Days),y=mean,group=RiverType,color=RiverType,shape=RiverType))+geom_line(size=1)+facet_wrap(~Treatment2)+ylab("Bacterial Shannon Diversity (+/- SEM)")+
  geom_point(data=ShannonData,aes(x=as.factor(Days),y=value))+ geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+geom_jitter()+xlab("Days")+theme(legend.justification=c(1,0), legend.position=c(.85,0.0),legend.background = element_rect(fill = "transparent"))
BacterialShannonPlot


ggplot(ShannonData,aes(x=SampleType,y=value,color=RiverType))+geom_point()+facet_wrap(~Treatment2)


```

Shannon diversity comparison between paired oak and chestnut samples
```{r}

PairedShannon<-ggpaired(ShannonData,x="Treatment2",y="value",id="Paired",color="Treatment",line.color = "gray", line.size = 0.4,)+ylab("Bacterial Shannon Diversity")+xlab("Leaf Type")+stat_compare_means(paired=T)+rremove("legend")+scale_color_manual(values=cbPalette)
PairedShannon



theme_set(theme_bw(base_size = 10)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme

dev.off()
tiff("Figures/BacterialDiversityFig.tiff", width = 174, height = 84, units = 'mm', res = 600)
ggarrange(BacterialShannonPlot,PairedShannon,
          labels = c("a", "b"),
          ncol = 2, nrow = 1)
dev.off()

theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme






CombinedDifference<-data.frame()
for (i in unique(ShannonData$Paired)){
  Subset<-subset(ShannonData, Paired==i)
  LeafQ<-subset(Subset,Treatment=="Q")
  LeafC<-subset(Subset,Treatment=="C")
  Difference<-LeafQ$value-LeafC$value
  CombinedDifference<-rbind(CombinedDifference,Difference)
}
colnames(CombinedDifference)<-"Difference"
hist(CombinedDifference$Difference)
mean(CombinedDifference$Difference)
sd(CombinedDifference$Difference)/sqrt(length(CombinedDifference$Difference))

ggplot(ObservedData,aes(x=Days,y=value,color=SampleType))+facet_wrap(RiverType~Treatment2)+geom_point()+ylab("Bacterial Richness (# ASV's)")



```



```{r}
plot_richness(physeq, x="Date",color="RiverType",shape="Treatment", measures=c("Observed"))+ylab("Observed Species")+geom_point()+scale_fill_manual(cbPalette)+geom_point(size=5)


Observed<-plot_richness(physeq, x="Date",color="RiverType",shape="Treatment", measures=c("Observed"))+ylab("Observed Species")+geom_point()+scale_fill_manual(cbPalette)+geom_point(size=5)+facet_wrap(~Site)
ObservedData<-Observed$data

ObservedAOV<-aov(value~RiverType,data=ObservedData)
summary(ObservedAOV)


ggplot(ObservedData,aes(x=Days,y=value,color=RiverType,shape=Treatment2))+geom_point(size=3)+ylab("Bacterial Richness (ASV)")



```


# Taxa plots

```{r,warning=F}
###########
#Phylum level plot
###########
df <- psmelt(PhylumLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Phylum","Treatment2","Days","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
PhylumPlotTreatment=ggplot(Trtdata, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Phylum),colour="black", stat="identity")+
  facet_grid(RiverType~Treatment2)+xlab("Days")+ylab("Relative Abundance (> 3%, SEM)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
PhylumPlotTreatment

dfCP<-subset(df,Treatment=="C"&RiverType=="Perennial")
compare_means(Abundance ~ Date, data = dfCP, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfCI<-subset(df,Treatment=="C"&RiverType=="Intermittent")
compare_means(Abundance ~ Date, data = dfCI, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns



dfQP<-subset(df,Treatment=="Q"&RiverType=="Perennial")
compare_means(Abundance ~ Date, data = dfQP, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota

dfQPActinobacteria<-subset(dfQP,Phylum=="Actinobacteriota")
compare_means(Abundance ~ Date, data = dfQPActinobacteria, group.by = "Phylum", p.adjust.method = "fdr",method="wilcox.test") # Actinobacteriota

#Actinobacteriota plot
TrtdataActino <- ddply(dfQPActinobacteria, c("Phylum","Days"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
ActinoPlot<-ggplot(TrtdataActino, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Phylum),colour="black", stat="identity")+xlab("Days")+ylab("Rel. Abundance Actinobacteriota (%, SEM)")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+theme(legend.position="none")

ActinoPlot



theme_set(theme_bw(base_size = 9)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme
PhylumPlotTreatment
ActinoPlot



dev.off()
tiff("Figures/BacterialTaxFigure.tiff", width = 174, height = 84, units = 'mm', res = 1200)
ggarrange(PhylumPlotTreatment,ActinoPlot,
          labels = c("a", "b"),
          ncol = 2, nrow = 1)
dev.off()

theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme





dfQI<-subset(df,Treatment=="Q"&RiverType1=="Intermittent")
compare_means(Abundance ~ Date, data = dfQI, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns


dfC<-subset(df,Treatment=="C")
compare_means(Abundance ~ Date, data = dfC, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns # Actinobacteria close


dfQ<-subset(df,Treatment=="Q")
compare_means(Abundance ~ Date, data = dfQ, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota



dfC<-subset(df,Treatment=="C")
compare_means(Abundance ~ RiverType1, data = dfC, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test")


dfQ<-subset(df,Treatment=="Q")
compare_means(Abundance ~ RiverType, data = dfQ, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test")


dfPer<-subset(df,RiverType=="Perennial")
compare_means(Abundance ~ Treatment, data = dfPer, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota significant, Bacteroidota nearly sig

ddply(dfPer, c("Phylum","Treatment","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N))



dfInt<-subset(df,RiverType=="Intermittent")
compare_means(Abundance ~ Treatment, data = dfInt, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota significant, Bacteroidota sig

ddply(dfInt, c("Phylum","Treatment","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N))







df <- psmelt(PhylumLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Phylum", "Treatment2","Days"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
#Trtdata
cdataplot=ggplot(Trtdata, aes(x=Days,y=mean))+geom_bar(aes(fill = Treatment2),colour="black", stat="identity")+ facet_grid(Phylum~Treatment2)+xlab("Days")+ylab("Relative Abundance (%, SEM)") + theme(axis.text.x = element_text(angle = 0, hjust = 0.5))+theme(legend.position = "none")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_manual(values=cbPalette)
cdataplot

```



Phylum comparisons between river type by date

```{r}
df <- psmelt(PhylumLevel)
dfQ21<-subset(df,Treatment=="Q"&Days=="21")
compare_means(Abundance ~ RiverType, data = dfQ21, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfQ42<-subset(df,Treatment=="Q"&Days=="42")
compare_means(Abundance ~ RiverType, data = dfQ42, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfQ63<-subset(df,Treatment=="Q"&Days=="63")
compare_means(Abundance ~ RiverType, data = dfQ63, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfQ84<-subset(df,Treatment=="Q"&Days=="84")
compare_means(Abundance ~ RiverType, data = dfQ84, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfQ105<-subset(df,Treatment=="Q"&Days=="105")
compare_means(Abundance ~ RiverType, data = dfQ105, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfQ126<-subset(df,Treatment=="Q"&Days=="126")
compare_means(Abundance ~ RiverType, data = dfQ126, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns






dfC21<-subset(df,Treatment=="C"&Days=="21")
compare_means(Abundance ~ RiverType, data = dfC21, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC42<-subset(df,Treatment=="C"&Days=="42")
compare_means(Abundance ~ RiverType, data = dfC42, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC63<-subset(df,Treatment=="C"&Days=="63")
compare_means(Abundance ~ RiverType, data = dfC63, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC84<-subset(df,Treatment=="C"&Days=="84")
compare_means(Abundance ~ RiverType, data = dfC84, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC105<-subset(df,Treatment=="C"&Days=="105")
compare_means(Abundance ~ RiverType, data = dfC105, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC126<-subset(df,Treatment=="C"&Days=="126")
compare_means(Abundance ~ RiverType, data = dfC126, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

dfC<-subset(df,Treatment=="C")
dfQ<-subset(df,Treatment=="Q")
compare_means(Abundance ~ RiverType, data = dfQ, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns
compare_means(Abundance ~ RiverType, data = dfC, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns


compare_means(Abundance ~ SampleType, data = dfQ, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns

compare_means(Abundance ~ SampleType, data = dfC, group.by = "Phylum", p.adjust.method = "fdr",method="kruskal.test") # ns


```


Family level bacterial communities

```{r}
df <- psmelt(FamilyLevel)
df$Abundance=df$Abundance*100
Trtdata <- ddply(df, c("Family","Treatment2","Days","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
FamilyPlotTreatment=ggplot(Trtdata, aes(x=Days,y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+
  facet_grid(RiverType~Treatment2)+xlab("Days")+ylab("Relative Abundance (> 3%, SEM)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
FamilyPlotTreatment

TrtdataC<-subset(Trtdata, Treatment2=="Chestnut")

FamilyPlotTreatmentC=ggplot(TrtdataC, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+
  facet_grid(RiverType~Family)+xlab("Days")+ylab("Relative Abundance (> 3%, SEM)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))
FamilyPlotTreatmentC



TrtdataQ<-subset(Trtdata, Treatment2=="Oak")

FamilyPlotTreatmentQ=ggplot(TrtdataQ, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+
  facet_grid(RiverType~Family)+xlab("Days")+ylab("Relative Abundance (> 3%, SEM)")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),legend.position="none")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))
FamilyPlotTreatmentQ



  


dfCP<-subset(df,Treatment=="C"&RiverType=="Perennial")
compare_means(Abundance ~ Date, data = dfCP, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") 

dfCPSphingo<-subset(dfCP,Family=="Sphingomonadaceae")
compare_means(Abundance ~ Date, data = dfCPSphingo, group.by = "Family", p.adjust.method = "fdr",method="wilcox.test")


TrtdataSphingo <- ddply(dfCPSphingo, c("Family","Days"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
SphingoPlot<-ggplot(TrtdataSphingo, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+xlab("Days")+ylab("Relative Abundance Sphingo (%, SEM)")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+theme(legend.position="none")

SphingoPlot



dfCI<-subset(df,Treatment=="C"&RiverType=="Intermittent")
compare_means(Abundance ~ Date, data = dfCI, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # ns



dfQP<-subset(df,Treatment=="Q"&RiverType=="Perennial")
compare_means(Abundance ~ Date, data = dfQP, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota

# dfQPActinobacteria<-subset(dfQP,Family=="Actinobacteriota")
# compare_means(Abundance ~ Date, data = dfQPActinobacteria, group.by = "Family", p.adjust.method = "fdr",method="wilcox.test") 









dfQI<-subset(df,Treatment=="Q"&RiverType=="Intermittent")
compare_means(Abundance ~ Date, data = dfQI, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # ns


dfC<-subset(df,Treatment=="C")
compare_means(Abundance ~ Date, data = dfC, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # ns # Actinobacteria close


dfQ<-subset(df,Treatment=="Q")
compare_means(Abundance ~ Date, data = dfQ, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota



dfC<-subset(df,Treatment=="C")
compare_means(Abundance ~ RiverType1, data = dfC, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")


dfQ<-subset(df,Treatment=="Q")
compare_means(Abundance ~ RiverType, data = dfQ, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test")


dfPer<-subset(df,RiverType=="Perennial")
compare_means(Abundance ~ Treatment, data = dfPer, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota significant, Bacteroidota nearly sig

ddply(dfPer, c("Family","Treatment","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N))



dfInt<-subset(df,RiverType=="Intermittent")
compare_means(Abundance ~ Treatment, data = dfInt, group.by = "Family", p.adjust.method = "fdr",method="kruskal.test") # Actinobacteriota significant, Bacteroidota sig

ddply(dfInt, c("Family","Treatment","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N))





Trtdata <- ddply(df, c("Family","Treatment2","Days","RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)





TrtdataFamilySubset<-subset(Trtdata,Family=="Flavobacteriaceae"|Family=="Comamonadaceae"|Family=="Sphingomonadaceae"|Family=="Spirosomaceae"|Family=="Methylophilaceae"|Family=="Rhodobacteraceae"|Family=="Rhodobacteraceae"|Family=="Chitinophagaceae")

FamilySubsetQ<-subset(TrtdataFamilySubset, Treatment2=="Oak")

FamilyPlotTreatmentSubsetQ=ggplot(FamilySubsetQ, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+
  facet_grid(RiverType~Family)+xlab("Days")+ylab("Relative Abundance Oak (%, SEM)")+
  theme(legend.position="none",axis.text.x = element_text(angle = 45, hjust = 1))+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))
FamilyPlotTreatmentSubsetQ

FamilySubsetC<-subset(TrtdataFamilySubset, Treatment2=="Chestnut")
FamilyPlotTreatmentSubsetC=ggplot(FamilySubsetC, aes(x=as.factor(Days),y=mean))+geom_bar(aes(fill = Family),colour="black", stat="identity")+
  facet_grid(RiverType~Family)+xlab("Days")+ylab("Relative Abundance Chestnut (%, SEM)")+
  theme(legend.position="none",axis.text.x = element_text(angle = 45, hjust = 1))+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))
FamilyPlotTreatmentSubsetC



theme_set(theme_bw(base_size = 12.5)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme


dev.off()
tiff("Figures/BacteriaFamilyLevelByDay.tiff", width = 275, height = 275, units = 'mm', res = 600)
ggarrange(FamilyPlotTreatmentSubsetQ,FamilyPlotTreatmentSubsetC,
          labels = c("a", "b"), nrow = 2)
dev.off()





```



# Genus level random forest predictions

## RiverType
```{r}
set.seed(155258)
GenusLevelRF = filter_taxa(GenusAll, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower than 1%

ForestData=GenusLevelRF#Change this one so you dont have to rewrite all variables
predictors=t(otu_table(ForestData))
response <- as.factor(sample_data(ForestData)$RiverType)
rf.data <- data.frame(response, predictors)
MozzieForest <- randomForest(response~., data = rf.data, ntree = 1000,importance=T)
print(MozzieForest)#returns overall Random Forest results



```



## Leaf type

```{r,warning=F}
set.seed(155258)
GenusLevelRF = filter_taxa(GenusAll, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower than 1%

ForestData=GenusLevelRF#Change this one so you dont have to rewrite all variables
predictors=t(otu_table(ForestData))
response <- as.factor(sample_data(ForestData)$Treatment)
rf.data <- data.frame(response, predictors)
MozzieForest <- randomForest(response~., data = rf.data, ntree = 1000,importance=T)
print(MozzieForest)#returns overall Random Forest results

imp <- importance(MozzieForest)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest testto classify treatment 
imp <- data.frame(predictors = rownames(imp), imp)
imp.sort <- arrange(imp, desc(MeanDecreaseAccuracy))
imp.sort$predictors <- factor(imp.sort$predictors, levels = imp.sort$predictors)
imp.20 <- imp.sort[1:20, ]

ggplot(imp.20, aes(x = predictors, y = MeanDecreaseAccuracy)) +
  geom_bar(stat = "identity", fill = "indianred") +
  coord_flip() +
  ggtitle("Most important genera for classifying  samples\n by Treatment")#\n in a string tells it to start a new line



destroyX = function(es) {
  f = es
  for (row in c(1:nrow(f))){ #for each column in dataframe
    if (startsWith(row.names(f)[row], "X") == TRUE)  { #if starts with 'X' ..
      row.names(f)[row] <- substr(row.names(f)[row], 2, 100) #get rid of it
    }
    assign(deparse(substitute(es)), f, inherits = TRUE)
  }
}

TopPredictors<-subset(imp.20,MeanDecreaseAccuracy> 5)
destroyX(TopPredictors)


otunames <- row.names(TopPredictors)
r <- rownames(tax_table(ForestData)) %in% otunames
otunames
PredictorTable<-kable(tax_table(ForestData)[r, ])#returns a list of the most important predictors for Random Forest Classification

GenusRandomForestSubset = subset_taxa(GenusAll, row.names(tax_table(GenusAll))%in% otunames)

df <- psmelt(GenusRandomForestSubset)
df$Abundance=df$Abundance*100

dfSorted<- df[with(df, order(Genus)), ]
dfSorted$FamilyGenus<-paste0(dfSorted$Family,": \n",dfSorted$Genus)
Trtdata <- ddply(dfSorted, c("FamilyGenus", "Treatment2"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
TrtdataSorted<- Trtdata[with(Trtdata, order(FamilyGenus)), ]
TrtdataSorted


cdataplot=ggplot(Trtdata, aes(x=Treatment2,y=mean))+geom_bar(aes(fill = Treatment2),colour="black", stat="identity")+xlab("Leaf Type ")+ylab("Relative Abundance (%)") + theme(axis.text.x = element_text(angle = 0, hjust = 0.5))+theme(axis.title.x=element_blank(),legend.position="none")+facet_wrap(~FamilyGenus,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))+scale_fill_manual(values=cbPalette)
cdataplot

Comparisons<-compare_means(Abundance ~ Treatment, data = dfSorted, group.by = "FamilyGenus", p.adjust.method = "fdr",method="kruskal.test")
Comparisons
#write.csv(Comparisons, "GenusLevelBacteiralRFIndicatorsLeafType.csv")



```

## River Type  within Chestnut

```{r}
set.seed(155258)
GenusLevelRF = filter_taxa(GenusAll, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower than 1%
ChestnutRF<-subset_samples(GenusLevelRF, Treatment2=="Chestnut")
ForestData=ChestnutRF#Change this one so you dont have to rewrite all variables
predictors=t(otu_table(ForestData))
response <- as.factor(sample_data(ForestData)$RiverType)
rf.data <- data.frame(response, predictors)
MozzieForest <- randomForest(response~., data = rf.data, ntree = 1000,importance=T)
print(MozzieForest)#returns overall Random Forest results


imp <- importance(MozzieForest)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest testto classify treatment 
imp <- data.frame(predictors = rownames(imp), imp)
imp.sort <- arrange(imp, desc(MeanDecreaseAccuracy))
imp.sort$predictors <- factor(imp.sort$predictors, levels = imp.sort$predictors)
imp.20 <- imp.sort[1:20, ]

ggplot(imp.20, aes(x = predictors, y = MeanDecreaseAccuracy)) +
  geom_bar(stat = "identity", fill = "indianred") +
  coord_flip() +
  ggtitle("Most important genera for classifying  samples\n by River type (Chestnut)")#\n in a string tells it to start a new line



destroyX = function(es) {
  f = es
  for (row in c(1:nrow(f))){ #for each column in dataframe
    if (startsWith(row.names(f)[row], "X") == TRUE)  { #if starts with 'X' ..
      row.names(f)[row] <- substr(row.names(f)[row], 2, 100) #get rid of it
    }
    assign(deparse(substitute(es)), f, inherits = TRUE)
  }
}

TopPredictors<-subset(imp.20,MeanDecreaseAccuracy> 5)
destroyX(TopPredictors)


otunames <- row.names(TopPredictors)
r <- rownames(tax_table(ForestData)) %in% otunames
PredictorTable<-kable(tax_table(ForestData)[r, ])#returns a list of the most important predictors for Random Forest Classification

GenusRandomForestSubset = subset_taxa(ForestData, row.names(tax_table(ForestData))%in% otunames)

df <- psmelt(GenusRandomForestSubset)
df$Abundance=df$Abundance*100

dfSorted<- df[with(df, order(Genus)), ]
dfSorted$FamilyGenus<-paste0(dfSorted$Family,": \n",dfSorted$Genus)
Trtdata <- ddply(dfSorted, c("FamilyGenus", "RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
TrtdataSorted<- Trtdata[with(Trtdata, order(FamilyGenus)), ]
TrtdataSorted


cdataplotChestnut=ggplot(Trtdata, aes(x=RiverType,y=mean))+geom_bar(aes(fill = RiverType),colour="black", stat="identity")+xlab("River type ")+ylab("Relative Abundance (%, Chestnut only)") + theme(axis.text.x = element_text(angle = 0, hjust = 0.5))+theme(axis.title.x=element_blank(),legend.position="none")+facet_wrap(~FamilyGenus,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))#+scale_fill_manual(values=cbPalette)
cdataplotChestnut


Comparisons<-compare_means(Abundance ~ RiverType, data = dfSorted, group.by = "FamilyGenus", p.adjust.method = "fdr",method="kruskal.test")
Comparisons


```


## River Type  within Oak

```{r}
set.seed(45433)
GenusLevelRF = filter_taxa(GenusAll, function(x) mean(x) > 1e-2, TRUE) #filter out any taxa lower than 1%
OakRF<-subset_samples(GenusLevelRF, Treatment2=="Oak")
ForestData=OakRF#Change this one so you dont have to rewrite all variables
predictors=t(otu_table(ForestData))
response <- as.factor(sample_data(ForestData)$RiverType)
rf.data <- data.frame(response, predictors)
MozzieForest <- randomForest(response~., data = rf.data, ntree = 1000,importance=T)
print(MozzieForest)#returns overall Random Forest results


imp <- importance(MozzieForest)#all the steps that are imp or imp. are building a dataframe that contains info about the taxa used by the Random Forest testto classify treatment 
imp <- data.frame(predictors = rownames(imp), imp)
imp.sort <- arrange(imp, desc(MeanDecreaseAccuracy))
imp.sort$predictors <- factor(imp.sort$predictors, levels = imp.sort$predictors)
imp.20 <- imp.sort[1:20, ]

ggplot(imp.20, aes(x = predictors, y = MeanDecreaseAccuracy)) +
  geom_bar(stat = "identity", fill = "indianred") +
  coord_flip() +
  ggtitle("Most important genera for classifying  samples\n by River type (Oak)")#\n in a string tells it to start a new line



destroyX = function(es) {
  f = es
  for (row in c(1:nrow(f))){ #for each column in dataframe
    if (startsWith(row.names(f)[row], "X") == TRUE)  { #if starts with 'X' ..
      row.names(f)[row] <- substr(row.names(f)[row], 2, 100) #get rid of it
    }
    assign(deparse(substitute(es)), f, inherits = TRUE)
  }
}

TopPredictors<-subset(imp.20,MeanDecreaseAccuracy> 5)
destroyX(TopPredictors)


otunames <- row.names(TopPredictors)
r <- rownames(tax_table(ForestData)) %in% otunames
PredictorTable<-kable(tax_table(ForestData)[r, ])#returns a list of the most important predictors for Random Forest Classification

GenusRandomForestSubset = subset_taxa(ForestData, row.names(tax_table(ForestData))%in% otunames)

df <- psmelt(GenusRandomForestSubset)
df$Abundance=df$Abundance*100

dfSorted<- df[with(df, order(Genus)), ]
dfSorted$FamilyGenus<-paste0(dfSorted$Family,": \n",dfSorted$Genus)
Trtdata <- ddply(dfSorted, c("FamilyGenus", "RiverType"), summarise,
                 N    = length(Abundance),
                 mean = mean(Abundance),
                 sd   = sd(Abundance),
                 se   = sd / sqrt(N)
)
TrtdataSorted<- Trtdata[with(Trtdata, order(FamilyGenus)), ]
TrtdataSorted


cdataplotOak=ggplot(Trtdata, aes(x=RiverType,y=mean))+geom_bar(aes(fill = RiverType),colour="black", stat="identity")+xlab("River type ")+ylab("Relative Abundance (%, Oak only)") + theme(axis.text.x = element_text(angle = 0, hjust = 0.5))+theme(axis.title.x=element_blank(),legend.position="none")+facet_wrap(~FamilyGenus,scales = "free_y")+geom_errorbar(aes(ymin=mean-se,ymax=mean+se))#+scale_fill_manual(values=cbPalette)
cdataplotOak


Comparisons<-compare_means(Abundance ~ RiverType, data = dfSorted, group.by = "FamilyGenus", p.adjust.method = "fdr",method="kruskal.test")
Comparisons

#####
# Combine oak and chestnut indicators
######
theme_set(theme_bw(base_size = 8)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme

dev.off()
tiff("Figures/OakChestnutRFIndicators.tiff", width = 190, height = 84, units = 'mm', res = 1200)
ggarrange(cdataplotChestnut,cdataplotOak,
          labels = c("a", "b"),
          ncol = 2, nrow = 1)
dev.off()

theme_set(theme_bw(base_size = 18)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())) #Set font size and change default plot theme






```

# Bacterial Beta diversity (dispersions)

```{r}
set.seed(12345)
GPdist=phyloseq::distance(physeq, "bray")
beta=betadisper(GPdist, sample_data(physeq)$Site)
permutest(beta)
boxplot(beta)



```

# Bacterial beta diversity (PERMANOVA)

```{r}
set.seed(12345)
ord=ordinate(physeq,"PCoA", "bray")

GPdist=phyloseq::distance(physeq, "bray")
adonis(GPdist ~ RiverType*Treatment*Site+Date, as(sample_data(physeq), "data.frame"))



sample_data(physeq)$Treatment<-sample_data(physeq)$Treatment2
ordplot2=plot_ordination(physeq, ord,"samples", color="Treatment",shape="Treatment")+geom_point(size=4)+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
ordplot2<-ordplot2+ stat_ellipse(type= "norm",geom = "polygon", alpha = 1/4, aes(fill = Treatment))+facet_wrap(~Days)+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())#+ theme(legend.justification=c(1,0), legend.position=c(1,0))
ordplot2


###########
# Pairwise compariosn between oak and chestnut leaves
#############

Day21<-subset_samples(physeq, Days=="21")
set.seed(12345)
GPdist=phyloseq::distance(Day21, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day21), "data.frame"))



Day42<-subset_samples(physeq, Days=="42")
set.seed(12345)
GPdist=phyloseq::distance(Day42, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day42), "data.frame"))




Day63<-subset_samples(physeq, Days=="63")
set.seed(12345)
GPdist=phyloseq::distance(Day63, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day63), "data.frame"))





Day84<-subset_samples(physeq, Days=="84")
set.seed(12345)
GPdist=phyloseq::distance(Day84, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day84), "data.frame"))


Day105<-subset_samples(physeq, Days=="105")
set.seed(12345)
GPdist=phyloseq::distance(Day105, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day105), "data.frame"))




Day126<-subset_samples(physeq, Days=="126")
set.seed(12345)
GPdist=phyloseq::distance(Day126, "bray")
adonis(GPdist ~ Treatment, as(sample_data(Day126), "data.frame"))






```


```{r}


ordplot3=plot_ordination(physeq, ord,"samples", color="RiverType")+geom_point(size=4)+scale_colour_manual(values=cbPalette)+scale_fill_manual(values=cbPalette)
ordplot3<-ordplot3+ stat_ellipse(type= "norm",geom = "polygon", alpha = 1/4, aes(fill = RiverType))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+facet_wrap(~Days)#+ theme(legend.justification=c(1,0), legend.position=c(1,0))
ordplot3



```
